---
title: "Mitch - Group Project 1"
author: "Mitch Whalen"
date: "2023-10-27"
output: html_document
---
## Setup
```{r}
knitr::opts_chunk$set(echo = TRUE)

# Set directories
traindir <- "/Users/mitchellwhalen/Library/CloudStorage/GoogleDrive-maw9byk@virginia.edu/My Drive/UVA Fall 2023/Statistical Modeling/In Class/Train Data"
sourcedir <-"/Users/mitchellwhalen/Library/CloudStorage/GoogleDrive-maw9byk@virginia.edu/My Drive/UVA Fall 2023/Statistical Modeling/In Class/R Code"

# load data
setwd(sourcedir)
source("AccidentInput.R")

# load libraries
library(ggplot2)
library(dplyr)
library(psych)
library(lattice)
library(here)
library(ggpubr)
library(ggfortify)
library(MASS)
library(lindia)
library(data.table)
library(plyr)
library(scales)
library(grid)
library(GGally)
library(car)
library(lmtest)
library(RColorBrewer)
library("gplots")
library(caret)
```

## Load and filter data to extreme accidents

Here we are loading the accident data, assigning labels, and then filtering to only extreme
accidents. Extreme accidents are all accidents with damages above the upper whisker

```{r}
acts <- file.inputl(traindir)

totacts <- combine.data(acts)

##Build a data frame with only extreme accidents for ACCDMG

dmgbox <-boxplot(totacts$ACCDMG)

ggplot(as.data.frame(totacts$ACCDMG), aes(x=totacts$ACCDMG)) + 
  geom_boxplot(col= "steelblue") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

xdmg <- totacts[totacts$ACCDMG > dmgbox$stats[5],]

#remove 9/11
xdmg <- xdmg[-181,]

## Remove duplicates from xdmg and call new data frame xdmgnd
xdmgnd <- xdmg[!(duplicated(xdmg[, c("INCDTNO", "YEAR", "MONTH", "DAY", "TIMEHR", "TIMEMIN")])),]

# Setup categorical variables
xdmgnd$Cause <- rep(NA, nrow(xdmgnd))

xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "M")] <- "M"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "T")] <- "T"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "S")] <- "S"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "H")] <- "H"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "E")] <- "E"

# This new variable, Cause, has to be a factor

xdmgnd$Cause <- factor(xdmgnd$Cause)

# Convert Type to a factor and give more meaningful labels
xdmgnd$Type <- factor(xdmgnd$TYPE, labels = c("Derailment", "HeadOn", "Rearend", "Side", "Raking", "BrokenTrain", "Hwy-Rail", "GradeX", "Obstruction", "Explosive", "Fire","Other","SeeNarrative"))

# Type of train
xdmgnd$TYPEQ <- as.numeric(xdmgnd$TYPEQ)

# Now convert to factor with meaningful labels
xdmgnd$TYPEQ <- factor(xdmgnd$TYPEQ, labels = c("Freight", "Passenger", "Commuter", "Work",  "Single", "CutofCars", "Yard", "Light", "Maint"))

```

## Generating Hypothesis based on Type of Train and Type of Accident

We looked at the frequency of the different types of trains and types of accidents in order to
discern any notable patterns.

```{r, echo=FALSE}
# Bar Graph of TYPEQ in xdmgnd
ggplot(as.data.frame(table(xdmgnd$TYPEQ)), aes(x = Var1, y= Freq)) +
  geom_bar(stat="identity",fill= "steelblue")+ 
  ggtitle("Accident Frequency by TypeQ (xdmgnd)") +
  labs(x = "Type of Train")+
  theme(axis.text.x = element_text(size = 8,  angle = 45))

# Bar Graph of Type in xdmgnd
ggplot(as.data.frame(table(xdmgnd$Type)), aes(x = Var1, y= Freq)) +
  geom_bar(stat="identity",fill= "steelblue")+ 
  ggtitle("Accident Frequency by Type (xdmgnd)") +
  labs(x = "Type of Accident")+
  theme(axis.text.x = element_text(size = 8,  angle = 45))
```
Clearly the most extreme accidents disproportionately involve freight trains and derailment.

```{r, echo=FALSE}
# Plot scaled (log) accident damage grouped by TypeQ using bwplot (from lattice package)
bwplot(
  TYPEQ~ log(ACCDMG+1), 
  main = "Box Plots of Log(Accident Damage)", 
  xlab = "log(Damage ($))", ylab = "Type of Train", data = xdmgnd
)

# Plot scaled (log) accident damage grouped by Type using bwplot (from lattice package)
bwplot(
  Type~ log(ACCDMG+1), 
  main = "Box Plots of Log(Accident Damage)", 
  xlab = "log(Damage ($))", ylab = "Type of Accident", data = xdmgnd
)
```
Freight, Commuter, & Passenger trains typically have the greatest damages, while the different
types of accidents don't seem to vary much in terms of accident damage (GradeX has a very small
sample size), however we can see an head on collisions seem to be especially damaging and
derailment accidents have high potential for damage as shown by their outliers.

```{r, echo=False}
# types of each type of train's accidents
ggplot(xdmgnd, aes(x=TYPEQ, fill=Type), reorder(Type)) +
  geom_bar(position="fill") +
  scale_fill_brewer(palette = "Spectral") + #https://r-graph-gallery.com/38-rcolorbrewers-palettes.html
  coord_flip()
```

The most apparent pattern to us was the difference in types of accidents between Freight and
Passenger/Commuter trains. AS you can see freight trains were typically involved in derailment
accidents, while passenger/commuter trains were far more frequently involved in Hwy_Rail
Accidents. NOTE: Moving forward we will group Passenger & Commuter trains since both serve a
similar purpose and carry passengers.

The last aspect we wanted to explore was train speed. Because different tyoes of trains may
operate in different fashions, we suspect that this could influence their speed and potential to
affect accidents.

```{r, echo=False}
# Looking for interactions between TRNSPD & ACCDMG by TYPEQ
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ TYPEQ)

# Looking for interactions between TRNSPD & ACCDMG by Type
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ Type)
```
There appears to be some interaction here with train speed. It appears there is at least a
slight interaction between speed and accident damage when it comes to freight trains and
commuter trains. There also appears to be an interaction with some accident types including
derailments, head-on, and side accidents.

# Hypothesis

Based on the visualizations, one can see that the Freight train accidents are far more likely to
result in extreme outcomes than accidents involving other types of trains. Most extreme freight
train accidents are derailments. This contrasts, passenger and commuter trains where a greater
proportion accidents are of the highway-rail type. Given these observations, I hypothesize that
the type of train (specifically the levels of freight or passenger/commuter), the type of
accident (specifically the levels of derailment or Hwy-Rail), and their interaction can be used
to predict accident damage. In addition, it is apparent that there is a significant interaction
between train speed and accident damage, and if we group by type of accident and type of train
we can see that the strength of the interaction appears to vary based on type of accident and type of train.

H0: Type of train (PassCom or Freight), type of accident (Derail or Hwy_Rail), and train speed
have no significant relationship with accident damage
Ha: Type of train (PassCom or Freight), type of accident (Derail or Hwy_Rail), and train speed
have a significant relationship with accident damage

## Analysis & Model Building

Prepping to test hypothesis:
```{r}
# Remove rows in dmgnd where there is a null value for TYPEQ or Type (got code from ChatGPT)
xdmgnd <- xdmgnd[complete.cases(xdmgnd[c("TYPEQ", "Type")]), ]

# Create Necessary Dummy Variables
xdmgnd$Derail <- (xdmgnd$Type == "Derailment")
xdmgnd$Hwy_Rail <- (xdmgnd$Type == "Hwy-Rail")
xdmgnd$Freight <- (xdmgnd$TYPEQ == "Freight")
xdmgnd$PassCom <- ifelse(xdmgnd$Type %in% c("Passenger", "Commuter"), TRUE, FALSE) #Grouped because commuter train carries passengers
```

Multiple Linear Regression using TYPEQ to Predict ACCDMG (LM1)
```{r}
xdmgnd.lm1<-lm(ACCDMG~Freight+PassCom, data=xdmgnd)
summary(xdmgnd.lm1)
```
Multiple Linear Regression using Type to Predict ACCDMG (LM2)
```{r}
xdmgnd.lm2<-lm(ACCDMG~Derail+Hwy_Rail, data=xdmgnd)
summary(xdmgnd.lm2)
```
Multiple Linear Regression using Type & TYPEQ to Predict ACCDMG (LM3)
```{r}
xdmgnd.lm3<-lm(ACCDMG~Freight+PassCom+Derail+Hwy_Rail,data=xdmgnd)
summary(xdmgnd.lm3)
```
Models LM1-3 have almost no predictive value based on their R^2. We will explore the
2nd order model as well as their interaction with speed to try and build a more predictive model

Multiple Linear Regression using Type & TYPEQ plus interactions to Predict ACCDMG (LM4)
```{r}
xdmgnd.lm4<-lm(ACCDMG~(Freight+PassCom+Derail+Hwy_Rail)^2,data=xdmgnd)
summary(xdmgnd.lm4)
```
Still very low predictive value

Will now add in train speed to hopefully improve model performance.
Multiple Linear Regression main effects using Type, TYPEQ, & TRNSPD to Predict ACCDMG (LM5)
```{r}
xdmgnd.lm5<-lm(ACCDMG~Freight+PassCom+Derail+Hwy_Rail+TRNSPD,data=xdmgnd)
summary(xdmgnd.lm5)
```
Multiple Linear Regression 2nd order model using Type, TYPEQ, & TRNSPD to Predict ACCDMG (LM6)
```{r}
xdmgnd.lm6 <- lm(ACCDMG ~ (Freight + PassCom + Derail + Hwy_Rail + TRNSPD)^2, data = xdmgnd)
summary(xdmgnd.lm6)
AIC(xdmgnd.lm6)
```
Adding TrnSpd to the model instantly boosted the model demonstrating its importance in
predicting accident damage. Including its interactions also improved the model slightly.

#Diagnostics, Transformations, & Variable Selection
We first needed to examine our diagnostic plots to see if LM6 is was meeting basic regression assumptions
```{r, echo=False}
autoplot(xdmgnd.lm6, which=1:6, label.size = 3) + theme_bw()
```
Immediately we can see clear heteroscedasticity as evidenced by the residual v fitted and QQ plot. We cna also, see at least 3 observations with concerning Cook's distance values.
Therefore we know we will need to transform our response variable. For this we will use a Box Cox transformation
```{r}
# Box Cox
gg_boxcox(xdmgnd.lm6)
#The best lambda and store in L
L<-boxcox(xdmgnd.lm6, plotit = F)$x[which.max(boxcox(xdmgnd.lm6, plotit = F)$y)] 
L
```
Now that we have an optimal lambda value of -0.5, we can use this to transform accident damage and improve LM6
```{r}
# The model with the best lamda transformation (LM6.boxcox)
xdmgnd.lm6.boxcox<-lm((ACCDMG^L-1)/L~(Freight + PassCom + Derail + Hwy_Rail + TRNSPD)^2,data=xdmgnd)
# Display regression results for boxcox model
summary(xdmgnd.lm6.boxcox)
AIC(xdmgnd.lm6.boxcox)
```
This transformation also greatly improved the predictive value of the model.

Now we will recheck the diagnostics plots
```{r,echo=False} 
autoplot(xdmgnd.lm6.boxcox, which=1:6, label.size = 3) + theme_bw()
```
As we can see we have addressed the galring heteroscedasticity issue. Some heteroscedasticity
remains as we can see in the residuals vs fitted plot, but we can see it is clearly much improved. Also, there are no concerning Cook's distance values.

We next aimed to trim our model using stepwise regression
We used backwards elimination to trim our model and remove insignificant predictors.
```{r}
xdmgnd.lm6.step <- step(xdmgnd.lm6.boxcox, direction = "backward", Trace = F)
summary(xdmgnd.lm6.step) 
AIC(xdmgnd.lm6.step)
anova(xdmgnd.lm6.boxcox,xdmgnd.lm6.step)
```
This backward elimination process removed the PassCom variable and its interactions as well as
Hwy_Rail:TRNSPD and FreightTRUE:Hwy_RailTRUE
The AIC is slightly lower than the full model but Adj. R^2 is the same.
Because the partial f-test has a p-value > 0.05, we can be confident that none of the removed
predictors were significant and we will move forward with the stepwise mode

## Evidence & Recommendation to FRA

COMING SOON